Results for working group papers

### 2016-03-28 catch_portfolios_mtg3.Rmd
updated Rmd file to include filtering example

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(MatchIt)
load("cfecCleaned.Rdata")
```

@@

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')
```

>>>>

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
load("cfecCleaned.Rdata")
```

@@

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')
```
# Paper 1:

To start, we can coarsely summarize the variability in total revenue by species through time, 
```{r, fig.pos="placeHere", fig.cap = "Log(total) revenue of each spp"}
spTot = group_by(cfec, specn, year) %>%
  summarize(totRev.bySpec = log(sum(g_earn, na.rm=T))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev.bySpec)) %>%   
  as.data.frame
ggplot(spTot, aes(x=year,y=totRev.bySpec,group=as.factor(specn))) +geom_line(aes(x=year,y=totRev.bySpec, color=as.factor(specn)))
```

```{r, fig.pos="placeHere", fig.cap = "CV of log(total revenue) across species (sd in log space)"}
spTot = group_by(spTot, year) %>%
  summarize(sd_yr = sd(totRev.bySpec, na.rm=T)) %>% 
  as.data.frame()
ggplot(spTot, aes(x=year,y=sd_yr)) + geom_point()

## Filtering out unique vessel - person combinations  
Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year. Some people might fish on multiple vessels within a year, and vessels might be used by multiple people within a year.

For each individual in each year, we can calculate the effective number of vessels used, and trends in that (mean across people).   
```{r}
# calculate diversity of individuals across vessels / time
cfec.boatDiv = group_by(cfec, p_holder, year) %>% 
  summarize(effBoats = 1 / sum((table(cadfg) / sum(table(cadfg)))^2) ) %>% 
  data.frame

cfec.boatDiv = as.data.frame(cfec.boatDiv)
g = gam(as.numeric(effBoats) ~ s(as.numeric(year)), data=cfec.boatDiv)
gpred = predict(g, newdata=list(year=1985:2014), se.fit=TRUE)
plot(1985:2014, gpred$fit, type="l", lwd=3, ylim=c(1.03,1.1))
lines(1985:2014, gpred$fit+2*gpred$se.fit)
lines(1985:2014, gpred$fit-2*gpred$se.fit)
```

Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.  
```{r}
cfecnew = cfec
cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
  summarize(nBoat = length(unique(cadfg))) %>%
  filter(nBoat==1) %>% 
  left_join(cfecnew)

# Deal with setnet fisheries, where lots of p_holder
# share same codes
cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
  summarize(nPerson = length(unique(p_holder))) %>%
  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
  left_join(cfec.1boat)

cfecnew = cfec.1boat
dim(cfecnew)[1]
```

So there's about 50% of the records remaining -- these are unique individial / vessel combinations. Side note: all of the set net / razor clam permits and individuals have been included. Not sure we want to drop them out [Milo - thoughts? ]

Now we can create a unique *id* column for individual - vessel combinations.

```{r}
cfecnew$id = paste(cfecnew$p_holder,cfecnew$cadfg)
# calculate number of unique person - vessel pairs ~ 40700
length(unique(cfecnew$id))
```

Just for demo purposes, we can start by plotting mean - CV relationships for each of these individual- vessel combinations (across years).

```{r, fig.pos="placeHere", fig.cap="Mean and variability (CV) of annual earnings (across time) for unique people-vessel combinations."}
meanVar = group_by(cfecnew, id, year) %>% 
  summarize(totRev = sum(g_earn)) %>% 
  as.data.frame() %>%
  group_by(id) %>%
  summarize(m = mean(log(totRev+1),na.rm=T), cv = sd(totRev,na.rm=T)/mean(totRev,na.rm=T))
plot(meanVar$m, meanVar$cv, xlab="Mean log(annual revenue)", ylab = "CV annual revenue")
```

One question we also were chatting about was what the distribution of the percent change metric looked like. 
```{r mergeYears}

totalcfec = group_by(cfecnew, id, year) %>% 
  summarize(totrev = sum(g_earn))

totalcfec$year = as.numeric(totalcfec$year)
dummyDF = expand.grid(id = unique(totalcfec$id), 
  year = min(totalcfec$year) : max(totalcfec$year), 
  stringsAsFactors = F)
# expand the data frame w/missing years
cfecExp = left_join(dummyDF, totalcfec) %>%
  arrange(id, year)
cfecExp = group_by(cfecExp, id) %>%
  mutate(difflog = c(NA, diff(log(totrev))), meanRev = c(NA, zoo::rollmean(totrev, 2))) %>%
  filter(!is.na(difflog))
```

This shows the pattern we'd hope to expect -- no trend as a function of mean revenue, . Namely that people making less than about 35K / year tend to have negative returns from one year to the next, and people with revenues > 35K -- especially the high earners -- tend to have positive returns. We also see decreased variability as a function of returns -- maybe because investments in technology, permits, etc.
```{r, fig.pos="placeHere", fig.cap = "Distribution of percent change in total revenue + relationship (or lack of) with total annual revenue"}
par(mfrow = c(2,1), mai=c(0.5,0.5,0.1,0.1))
plot(log(cfecExp$meanRev), cfecExp$difflog, xlab = "log(revenue)", ylab="percent change in revenue", cex=0.7)
abline(0, 0, col="red")

hist(cfecExp$difflog, 80, xlab="Percent return (empirical)", col="grey", xlim= quantile(cfecExp$difflog, c(.005,0.995)), main="")
